---
title: "Analysis of Exp io RT-qPCR Data: Average Ct Calculations"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath('~/Documents/Master/PCR/RStudio/RStudio_qPCR_Exp1.cDNA/Data Analysis'))

```


Load files into R.

```{r load files into notebook}

# Load the saved list of data frames (dfs) into R Notebook
  load("/Users/elinberntsson/Documents/Master/PCR/RStudio/RStudio_qPCR_Exp1.cDNA/all_filtered.RData")

# Dfs we have to work with
  lapply(names(complete_filtered_list), function(x) assign(x, complete_filtered_list[[x]], envir = .GlobalEnv))

```

Look through dfs and check for outliers in the Ct values.
Remove any outliers.

```{r removal of any outliers}

# Go through dfs and remove if any outliers in the Ct column

  #Filtered_d13.2 <- Filtered_d13.2[Filtered_d13.2$Ct != 'Undetermined',] # Have not checked more outliers than the rows with 'Undetermined' in them - will do this!

# Replace 'Undetermined' with 'NA' instead

  # replaced_ct <- as.numeric(replace(Filtered_d13.2$Ct, Filtered_d13.2$Ct == 'Undetermined', NA))
  # 
  # Filtered_d13.2 <- cbind(Filtered_d13.2[, !colnames(Filtered_d13.2) %in% 'Ct'], Ct = replaced_ct)

# D0 and D13.2 - Undetermined to NA
  Filtered_d0$Ct <- ifelse(Filtered_d0$Ct == "Undetermined", NA, as.numeric(Filtered_d0$Ct))

  Filtered_d13.2$Ct <- ifelse(Filtered_d13.2$Ct == "Undetermined", NA, as.numeric(Filtered_d13.2$Ct))
  
# D3.1 and D8.1 SPINK2 outliers - Ct outliers to NA
  
  # D3.1 SPINK2 3uM Position E6 possible outlier
  Filtered_d3.1$Ct[Filtered_d3.1$Position == "E6"] <- NA 
  
  # D8.1 SPINK2 1.8uM Position E3 possible outlier
  Filtered_d8.1$Ct[Filtered_d8.1$Position == "E3"] <- NA
  
  # D8.1 SPINK2 5uM Position E7 possible outlier
  Filtered_d8.1$Ct[Filtered_d8.1$Position == "E7"] <- NA

```


Calculate mean Ct values for gene markers within each df.

```{r calculate mean Ct for filtered dfs}

# Function that calculates mean of every n:th row in column with Ct values

  mean_Ct <- function(df, column) {
    n <- 3 # 3 technical replicates we want to average over
  
    means_vector <- numeric()
  
    # Calculate means for each set of n rows
    for (i in seq(1, nrow(df), by = n)) {
    
      end_i <- min(i + n -1, nrow(df))
      means_vector <- c(means_vector, mean(as.numeric(df[i:end_i, column]), na.rm = TRUE))
    }
  
    return(means_vector) # Includes all calculated Ct means
  }


# Calculate mean Ct for genes in D: 0, 3, 8, 13 dfs with function created above

  gene_rows_d0 <- seq(from = 1, by = 3, length.out = 16) # To get row information for each gene and condition from day 0
  gene_rows <- seq(from = 1, by = 3, length.out = 24) # To get row information for each gene and condition from day: 3, 8, 13
  
  #Day 0
  Mean_d0 <- cbind(Filtered_d0[gene_rows_d0, !colnames(Filtered_d0) %in% 'Ct'], 'Mean Ct' = mean_Ct(Filtered_d0, 'Ct'))
  
  #Day 3
  Mean_d3.1 <- cbind(Filtered_d3.1[gene_rows, !colnames(Filtered_d3.1) %in% 'Ct'], 'Mean Ct' = mean_Ct(Filtered_d3.1, 'Ct'))
  Mean_d3.2 <- cbind(Filtered_d3.2[gene_rows, !colnames(Filtered_d3.2) %in% 'Ct'], 'Mean Ct' = mean_Ct(Filtered_d3.2, 'Ct'))
  
  #Day 8
  Mean_d8.1 <- cbind(Filtered_d8.1[gene_rows, !colnames(Filtered_d8.1) %in% 'Ct'], 'Mean Ct' = mean_Ct(Filtered_d8.1, 'Ct'))
  Mean_d8.2 <- cbind(Filtered_d8.2[gene_rows, !colnames(Filtered_d8.2) %in% 'Ct'], 'Mean Ct' = mean_Ct(Filtered_d8.2, 'Ct'))
  
  #Day 13
  Mean_d13.1 <- cbind(Filtered_d13.1[gene_rows, !colnames(Filtered_d13.1) %in% 'Ct'], 'Mean Ct' = mean_Ct(Filtered_d13.1, 'Ct'))
  Mean_d13.2 <- cbind(Filtered_d13.2[gene_rows, !colnames(Filtered_d13.2) %in% 'Ct'], 'Mean Ct' = mean_Ct(Filtered_d13.2, 'Ct'))
  


```

Calculate mean Ct values for the housekeeping gene markers.

```{r calculate average of ACTB and GAPDH Ct values to use as reference}

# Need to average the housekeeping gene Ct values

# mean_hk_3_CG<-mean(c(Mean_d3.1[Mean_d3.1$Gene=="Actb-SYBR" & Mean_d3.1$Condition=="3", "Mean Ct"], Mean_d3.1[Mean_d3.1$Gene=="GAPDH-Syber" & Mean_d3.1$Condition=="3", "Mean Ct"]))


## Day 0

mean_values_d0 <- c(mean(c(Mean_d0[Mean_d0$Gene == "Actb-SYBR" & Mean_d0$Bioreplica == '2.1', "Mean Ct"],
                          Mean_d0[Mean_d0$Gene == "GAPDH-Syber" & Mean_d0$Bioreplica == '2.1', "Mean Ct"])), 
                   mean(c(Mean_d0[Mean_d0$Gene == "Actb-SYBR" & Mean_d0$Bioreplica == '2.2', "Mean Ct"],
                          Mean_d0[Mean_d0$Gene == "GAPDH-Syber" & Mean_d0$Bioreplica == '2.2', "Mean Ct"])))

hk_full_d0 <- cbind(Mean_d0[1:2,1:4], Position = c('NA', 'NA'), Gene = c('ACTB_GAPDH', 'ACTB_GAPDH'), 'Mean Ct' = mean_values_d0)

full_hk_d0 <- rbind(hk_full_d0, Mean_d0[Mean_d0$Gene != 'Actb-SYBR' & Mean_d0$Gene != 'GAPDH-Syber',])



## Day 3, 8, 13


# Function that calculates mean Ct value for reference genes per condition (1.8, 3, 5 uM) and fills vector, named accordingly, with resulting values
  mean_hk <- function(df, day, replica, Condition) {
  
    mean_values <- numeric(length(Condition)) # Initialize a vector to store mean values
  
    for (i in 1:length(Condition)) {
    
        mean_value <- mean(c(df[df$Gene == "Actb-SYBR" & df$Condition == as.character(Condition[i]), "Mean Ct"],
                          df[df$Gene == "GAPDH-Syber" & df$Condition == as.character(Condition[i]), "Mean Ct"]))
        mean_values[i] <- mean_value # Store mean value in the vector
    }
  
    vector_name <- paste("mean_hk", '_', day, '.', replica, sep = "") # Create a name for the vector
    assign(vector_name, mean_values, envir = .GlobalEnv) # Assign the vector to a name in the global environment
  }


# Use mean_hk function

  Condition <- c(3, 1.8, 5) #Same for all D: 3, 8, 13
  
  #Day 3
  mean_hk(Mean_d3.1, 'd3', '1', Condition)
  mean_hk(Mean_d3.2, 'd3', '2', Condition)
  
  #Day 8
  mean_hk(Mean_d8.1, 'd8', '1', Condition)
  mean_hk(Mean_d8.2, 'd8', '2', Condition)
  
  #Day 13
  mean_hk(Mean_d13.1, 'd13', '1', Condition)
  mean_hk(Mean_d13.2, 'd13', '2', Condition)



# Create updated dfs with previous ACTB and GAPDH rows excluded, with new reference gene Ct values added in their place

hk_common_info <- Mean_d3.1[1:3,1:5] # Use Mean_d3.1 as basis, as all dfs (D: 3, 8, 13) look the same 
NA_position <- rep('NA', 3)
hk_name <- rep('ACTB_GAPDH', 3) # Combined reference gene name, same for all dfs


# hk_full: function that adds common info about the hk genes to the calculated mean Ct values per condition
## also exchanges the separate ACTB and GAPDH rows with newly generated ones

hk_full <- function(Mean_df, mean_hk_vector, day, replica) {
  
  # cbind info with calculated averages 
  hk_full <- cbind(hk_common_info, Position = NA_position, Gene = hk_name, 'Mean Ct' = mean_hk_vector)
  
  # rbind new reference gene dfs (hk_full) with original Mean_dX df
  full_hk <- rbind(hk_full, Mean_df[Mean_df$Gene != 'Actb-SYBR' & Mean_df$Gene != 'GAPDH-Syber',])
  
  full_hk_name <- paste("full_hk", '_', day, '.', replica, sep = "") # Create a name for the df with average hk Ct:s
  
  # Assign new full_hk df a name and save in .GlobalEnv
  assign(full_hk_name, full_hk, envir = .GlobalEnv) # Assign the vector to a name in the global environment
  
}


# Use hk_full function

  #Day 3
  hk_full(Mean_d3.1, mean_hk_d3.1, 'd3', '1')
  hk_full(Mean_d3.2, mean_hk_d3.2, 'd3', '2')
  
  #Day 8
  hk_full(Mean_d8.1, mean_hk_d8.1, 'd8', '1')
  hk_full(Mean_d8.2, mean_hk_d8.2, 'd8', '2')
  
  #Day 13
  hk_full(Mean_d13.1, mean_hk_d13.1, 'd13', '1')
  hk_full(Mean_d13.2, mean_hk_d13.2, 'd13', '2')
  
  


```


Calculate mean Ct values for housekeeping gene markers within the BPT datset.
```{r average hk genes in chipsc22 df}

  mean_hk_chipsc <- mean(Filtered_chipsc22[c(1, 2), 'Mean Ct']) # Average Ct value for reference gene from chipsc22 df

  hk_chipsc <- data.frame(Filtered_chipsc22[1,1:2], 'NA', 'ACTB_GAPDH', mean_hk_chipsc) # New row with new mean Ct

  # Exchange old hk gene values with newly calculated one mean_hk_chipsc
  names(hk_chipsc) <- names(Filtered_chipsc22)
  full_hk_chipsc <- rbind(hk_chipsc, Filtered_chipsc22[-c(1,2),])

```


Create a df with a column for each gene, including all calculated mean Ct values.
Create a cvs file with all data included.

```{r create dataframe with one column for each gene, with all mean Ct values included}

# We need one column with hk gene Ct values
# We need columns with gene of interest values

  
  
# for loop to create vectors with mean Ct values for each gene, to be merged with cbind() into combined_gene_analysis below
  genes_vector <- c('ACTB_GAPDH', 'SOX17-Sybr', 'RSPO3-Sybr', 'SPINK2-Sybr', 'HOXA9-Sybr', 'Runx1', 'Gata2-Sybr')
  gene_names <- c('hk', 'sox17', 'rspo3', 'spink2', 'hoxa9', 'runx1', 'gata2')
  
  gene_analysis_names <- NULL
  
  for (i in 1:length(genes_vector)) {
    gene_analysis_name <- paste(gene_names[i], 'gene', 'analysis', sep = '_')
    gene_analysis_names[[i]] <- gene_analysis_name
    
    # ct_per_gene <- c(full_hk_chipsc[full_hk_chipsc$Gene == genes_vector[i], 'Mean Ct'], 
    #                   order_hk_d3.1[order_hk_d3.1$Gene == genes_vector[i], 'Mean Ct'], 
    #                   order_hk_d3.2[order_hk_d3.2$Gene == genes_vector[i], 'Mean Ct'],
    #                   order_hk_d8.1[order_hk_d8.1$Gene == genes_vector[i], 'Mean Ct'],
    #                   order_hk_d8.2[order_hk_d8.2$Gene == genes_vector[i], 'Mean Ct'],
    #                   order_hk_d13.1[order_hk_d13.1$Gene == genes_vector[i], 'Mean Ct'],
    #                   order_hk_d13.2[order_hk_d13.2$Gene == genes_vector[i], 'Mean Ct'])
    
    ct_per_gene <- c(full_hk_chipsc[full_hk_chipsc$Gene == genes_vector[i], 'Mean Ct'],
                      full_hk_d0[full_hk_d0$Gene == genes_vector[i], 'Mean Ct'],
                      full_hk_d3.1[full_hk_d3.1$Gene == genes_vector[i], 'Mean Ct'],
                      full_hk_d3.2[full_hk_d3.2$Gene == genes_vector[i], 'Mean Ct'],
                      full_hk_d8.1[full_hk_d8.1$Gene == genes_vector[i], 'Mean Ct'],
                      full_hk_d8.2[full_hk_d8.2$Gene == genes_vector[i], 'Mean Ct'],
                      full_hk_d13.1[full_hk_d13.1$Gene == genes_vector[i], 'Mean Ct'],
                      full_hk_d13.2[full_hk_d13.2$Gene == genes_vector[i], 'Mean Ct'])
    
    assign(gene_analysis_name, ct_per_gene)
  }
  
  
  
  
group_var <- c('chiPSC22', rep(c('D0 2.1', 'D0 2.2')),
                          rep(c('D3 3uM', 'D3 1.8uM', 'D3 5uM'), 2), 
                          rep(c('D8 3uM', 'D8 1.8uM', 'D8 5uM'), 2),
                          rep(c('D13 3uM', 'D13 1.8uM', 'D13 5uM'), 2))

combined_gene_analysis <- data.frame(ACTB_GAPDH = as.numeric(hk_gene_analysis), 
                                     SOX17 = as.numeric(sox17_gene_analysis), 
                                     RSPO3 = as.numeric(rspo3_gene_analysis),
                                     SPINK2 = as.numeric(spink2_gene_analysis),
                                     HOXA9 = as.numeric(hoxa9_gene_analysis),
                                     RUNX1 = as.numeric(runx1_gene_analysis),
                                     GATA2 = as.numeric(gata2_gene_analysis),
                                     Group = group_var)

# Reorder data so all 1.8uM comes first, then 3uM...

combined_gene_analysis$Group <- factor(combined_gene_analysis$Group, levels = c('chiPSC22', 'D0 2.1', 'D0 2.2', 
                                                                                'D3 1.8uM', 'D8 1.8uM', 'D13 1.8uM', 
                                                                                'D3 3uM', 'D8 3uM', 'D13 3uM',
                                                                                'D3 5uM', 'D8 5uM', 'D13 5uM'))


all_genes_ordered <- combined_gene_analysis[order(combined_gene_analysis$Group),]


 write.csv(all_genes_ordered, file="genes_all_final.csv")
 
```





