---
title: "Sorting of Exp IO qPCR Data"
output:
  pdf_document: default
  html_document:
    self_contained: yes
    output_file: ElinHBerntsson_MasterThesis_DataSorting
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath('~/Documents/Master/PCR/RStudio/RStudio_qPCR_Exp1.cDNA'))

```



##### qPCR Data: Frozen cDNA chiPSC22 and Exp1 (D:0,3,8,13)


In this first part, R reads data into a common list and then extracts each list data frame (df) into separate dfs with names accordingly.
We also let R read the files of differing structure and content into a special list.

file_names is a list of all the files we want to load into R. 
list.files reads files in folder in alphabetical order - will need to adjust according to this order in the rest of the code e.g. in the day vector, etc.

read.delim2 reads all text files defined in code to be of interest into data_list. Skipping the first 10 rows to get correct columns names, and other functions.

```{r load files into R}

# Start by pointing R in the right direction; where to find the desired files.

  # Define the path to the directory containing the files, instead of setwd(), which can create difficulties 
  path_to_files <- "/Users/elinberntsson/Documents/Master/PCR/RStudio/RStudio_qPCR_Exp1.cDNA/Raw_Data_IO_Samples/"

  # List all files in the directory, list.files order files alphabetically by default
  file_names <- list.files(path = path_to_files)

  # Print the list of file names to see which files are included
  cat("Files inlcuded in list:\n", file_names)


  # Attach the path to the all the file names so R knows which directory to look in for files
  ## Made one for the files of same structure 
  full_file_paths_common <- paste0(path_to_files, file_names[2:7]) # You'll have to change file_names vector accordingly, if not all files are of same structure 
  ## One path for the special ones, e.g. chiPSC22 and D0/chiPSC22 data
  full_file_paths_special <- paste0(path_to_files, file_names[c(1,8)])


#sapply(full_file_paths, file.exists) # If you want to check if files in file_names exist in file path



# We want R to read all text files into a list of data frames (dfs) -> then extract each list element into separate dfs

  # Read text files defined in full_file_paths into a list of dfs with read.delim2. 
  ## Also e.g. skip 10 first rows to get correct column names, use headers as column names, etc.
  
  # One 'common'
  data_list_common <- lapply(full_file_paths_common, read.delim2, colClasses = c(Well="NULL"), header = TRUE, skip = 10, fill = TRUE)
  # One 'special'
  data_list_special <- lapply(full_file_paths_special, read.delim2, colClasses = c(Well="NULL"), header = TRUE, skip = 10, fill = TRUE)

```

<br>
<br>

In the 4 following chunks, the data from the data_list_common will be operated on. After this, the data with diverging structure will be handled. 


### Exp1 D:3,8,13

Each list element is extracted into separate dfs, with corresponding names. head() is used to demonstrate the structure of the complete, unsorted data.

```{r list elements extraction into dfs}

# Extract list elements into separate data frames, naming the dfs accordingly

  day_index <- rep(c('d3', 'd8', 'd13'), each = 2) # Add days here when have data for them - 2 of each because 2 bio.replicas
  replica <- rep(c(1,2), length(data_list_common)) # Create vector with repeated bio.replica number, as we have two replicas per day (& condition)
  
  # Store the names of the full dfs for later use
  full_df_names <- vector("list", length(data_list_common))

  # Make sure that the length of 'day' is the same as the length of 'data_list' - as a checkpoint that correct data --> correct df with corresponding name
  if(length(day_index) == length(data_list_common)) {
    
    for (i in 1:(length(data_list_common))) {
      
      full_df_name <- paste("full", '_', day_index[i], '.', replica[i], sep = "")
      full_df_names[[i]] <- full_df_name
      assign(full_df_name, data.frame(data_list_common[[i]])) # Creates df from list for each day, naming of df is dynamic
    
      # Print a message indicating the name of the data frame
      cat("Head of the data frame", full_df_name, ":\n")

      # Print the head of the newly created data frame
      print(head(get(full_df_name)))

    }
  }

```

<br>
<br>

In the following chunk of code the complete dfs are cleaned and sorted. 
Some rows with unnecessary information are excluded, data is ordered after gene markers (according to how they are structured on the plate).  

The cleaned and sorted data is then stored in a list of dfs (geneorder_list_common).

```{r clean and order after genes}

# Remove rows with unwanted data

  # Created vector with row numbers that we want to skip to clean data
  skiprow_d <- c(13:16, (13:16)+16, (13:16)+16*2, (13:16)+16*3, (13:16)+16*4, (13:16)+16*5, (13:16)+16*6, (13:16)+16*7, 129) #Looks like this because lazy

  # Store the names of the full dfs for later use
  clean_names <- vector("list", length(data_list_common))
  
  
  # Utilize skiprow_d vector to exclude rows we do not want
  ## Make sure that the length of 'day' is the same as the length of 'data_list_common' - as a checkpoint step
  
  if(length(day_index) == length(data_list_common)) {
    
    for (i in 1:length(data_list_common)) {
      
      full_df_name <- full_df_names[[i]]
  clean_name <- paste("clean", '_', day_index[i], '.', replica[i], sep = "")
  clean_names[[i]] <- clean_name
      assign(clean_name, get(full_df_name)[-skiprow_d,]) # Creates df from list for each day, naming of df is dynamic
      
    }
  }


# Order data frames after desired gene marker order

  # First, convert the 'gene' column to a factor with levels in the desired order: Use d3.1 as basis
  clean_d3.1$Detector.Name <- factor(clean_d3.1$Detector.Name, 
                                     levels = c('Actb-SYBR', 'GAPDH-Syber', 'SOX17-Sybr', 'RSPO3-Sybr', 'SPINK2-Sybr', 'HOXA9-Sybr', 'Runx1', 'Gata2-Sybr'))


  ## Create empty list to keep df:s in
  geneorder_list_common <- NULL

    for (i in 1:length(data_list_common)) {
      
  geneord_name <- paste("geneord", '_', day_index[i], '.', replica[i], sep = "")
  clean_name <- clean_names[[i]]
      assign(geneord_name, get(clean_name)[order(clean_d3.1$Detector.Name),]) # Creates df from list for each day, naming of df is dynamic
    
    # Fill list, use later when cbind() extra columns to dfs
    geneorder_list_common[[i]] <- get(geneord_name)
    
    }

```

<br>
<br>

In the following chunk we want to add columns with information about the experiment, conditions, etc. 

A df with desired columns are first created, and named accordingly. 
This column is then bound to the ordered dfs via cbind() command. A list of the new dfs is created (cbind_list).

```{r columns with info added}

# Create columns with important data (Exp_name, Time_point, Condition, Replicate, Temp.free_control), which we then want to add to correct dfs

  Exp1 <- rep('Exp1', 96) # Change accordingly 
  Condition <- rep(c(3, 'NA', 'NA', 'NA', 3, 3, 1.8, 1.8, 1.8, 5, 5, 5), 8) #mM - Strange order because of how sample data is loaded into R (as characters)
  Replicate1 <- rep(1, 96)
  Replicate2 <- rep(2, 96)
  Tempfree_control <- rep(c('no', 'yes', 'yes', 'yes', 'no', 'no', 'no', 'no', 'no', 'no', 'no', 'no'), 8) # Same goes for order here as with condition vector.


  # Timepoints changes depending on harvesing day, need to create these in e.g. a for loop like below
  ## Create vector with harvesing days to choose from
  all_days <- c('Day 3', 'Day 8', 'Day 13') # Add more days when have data, in order according to files loaded into R
  
  
  # Initialize time points and columns lists
  timepoint_list <- NULL
  columns_list <- NULL 


    for (i in 1:length(day_index)) {
      # Create variable names
      timepoint_names <- paste("Timepoint", "_", day_index[i], sep = "")
    
      # Match the day in 'all_days' with the corresponding 'day' entry 
      ## which() returns index of match if there is one, otherwise returns an empty vector of length = 0
      day_match <- which(gsub("Day ", "d", all_days) == day_index[i]) #gsub() replaced 'Day ' with 'd' so that matching can be done
      
      # Check if a matching day was found
      if(length(day_match) > 0){
        
        timepoint_values <- rep(all_days[day_match], 96) # If matching day, create vector with specific day that matches, e.g. 'Day 3' for 'd3'
        assign(timepoint_names, timepoint_values) # Assign the newly created time point vector to the created time point name
      
        # Append to list with time points dfs
        ## Assigns right timepoint_values to the correct list position given by timepoint_names
        timepoint_list[[timepoint_names]] <- timepoint_values
      
        # Create columns with all important data
        columns <- paste("columns", '_', day_index[i], '.', replica[i], sep = "")
        assign(columns, data.frame(Exp1, get(timepoint_names), Condition, replica[i], Tempfree_control))
      
        columns_list[[i]] <- get(columns)
        
      }
    }


  # Add created columns to data frames accordingly

  cbind_list_common <- NULL

  for (i in 1:length(day_index)) {
    
    cbind_df <- paste("cbound", '_', day_index[i], '.', replica[i], sep = "")
    assign(cbind_df, cbind(as.data.frame(columns_list[[i]]), as.data.frame(geneorder_list_common[[i]])))
  
    cbind_list_common[[i]] <- get(cbind_df)
    
  }

```

<br>
<br>

Here, columns containing information that is of no interest, and the the data (rows) from the negative controls are excluded.
The list of dfs we are left with contain all necessary data for continued analysis, utilizing whatever R package available. 

The data might need to be modified further to fit the required structure for whatever R package you want to use. 

```{r filtering of data}
  
# Create dfs with only columns of interest

  sorted_list_common <- NULL

  col_index <- c(1:7, 10)
  for (i in 1:length(day_index)) {
    sorted_data <- paste("sorted", '_', day_index[i], '.', replica[i], sep = '')
    assign(sorted_data, cbind_list_common[[i]][,col_index])
  
    sorted_list_common[[i]] <- get(sorted_data)
  }
  
    
# Remove template free control data
  
  Filtered_list_common <- NULL

  for (i in 1:length(day_index)) {
    
    # Create a name for the filtered data
    filtered_data_name <- paste('Filtered', '_', day_index[i], '.', replica[i], sep = '')
  
    # Directly filter the df where Tempfree_control is not equal to 'yes'
    filtered_data <- sorted_list_common[[i]][sorted_list_common[[i]]$Tempfree_control != 'yes',]
  
    # Assign a new names to the 1-8th columns
    col_names <- c('Experiment', 'Time point', 'Condition', 'BioReplica', 'Temp.free control', 'Position', 'Gene', 'Ct')
    names(filtered_data)[1:8] <- col_names

    # Assign the filtered data to a new variable in the environment
    assign(filtered_data_name, filtered_data)
    
    Filtered_list_common[[i]] <- get(filtered_data_name)
    
  }
  
  filtered_data <- NULL # To clear the last saved iteration data from loop
  
# Use head() to demonstrate how data looks
# print(head(Filtered_d3.1))

```

```{r hidden code for head(), echo=FALSE, results='asis'}

# To make the html file look nice: this chunk is added to print the full data, while not showing the actual code in the html file. 
  
  cat("Head of the data frame Filtered_d3.1 as an example:\n")
  cat(capture.output(head(Filtered_d3.1)), sep="\n")

```

<br>

In the following parts, the 'special' data files are handled and cleaned separately. First data from Exp 1 D0, followed by chiPSC22 data. 

Code starts with creating dfs with desired data - desired list elements from data_list_special are extracted into separate dfs, with corresponding names. 
head() is used to demonstrate the structure of the complete, unsorted data.
<br>

### Exp1 D0

The D0 data is handled similarly to the data from Day 3, 8 and 13 in the chunks above.

```{r create df - special data D0}

# Df with Exp1 D0 data is created 

# D0/chiPSC22
  assign(paste("full", '_', 'D0_chiPSC22', sep = ""), data.frame(data_list_special[[2]]))

# Use head() to demonstrate how data looks
# print(head(full_chiPSC22))

```

```{r hidden code for head() - special data D0, echo=FALSE, results='asis'}

# To make the html file look nice: this chunk is added to print the full data, while not showing the actual code in the html file. 
  
  cat("Head of the data frame full_D0/chiPSC22:\n")
  cat(capture.output(head(full_D0_chiPSC22)), sep="\n")

```

<br>

In following chunk of code, the data is cleaned. 
Some rows with unnecessary information are excluded. 

```{r clean data from unwanted rows - special data D0}

# Remove rows with unwanted data

  # Created vectors with row numbers that I want to skip to clean data
  skiprow_d <- c(13:16, (13:16)+16, (13:16)+16*2, (13:16)+16*3, (13:16)+16*4, (13:16)+16*5, (13:16)+16*6, (13:16)+16*7, 129) # For D0/chiPSC22 data


  # Utilize skiprow vector to exclude rows we do not want
  # D0/chiPSC22
  clean_D0_chipsc22 <- full_D0_chiPSC22[-skiprow_d,]
  
  
# Extract D0 data from the D0/chiPSC22 df to sort seperately
  chipsc_rows_all <- c(10:12, 22:24, 34:36, 46:48, 58:60, 70:72, 82:84, 94:96)
  clean_d0 <- clean_D0_chipsc22[-chipsc_rows_all,]
  
```

<br>
<br>

Here, data is ordered after gene markers (according to how they are structured on the plate). 

```{r order after genes - special data D0}

# Order df after desired gene marker order

  # First, convert the 'gene' column to a factor with levels in the desired order
  clean_d0$Detector.Name <- factor(clean_d0$Detector.Name, 
                                     levels = c('Actb-SYBR', 'GAPDH-Syber', 'SOX17-Sybr', 'RSPO3-Sybr', 'SPINK2-Sybr', 'HOXA9-Sybr', 'Runx1', 'Gata2-Sybr'))

  # Order after the desired order
  geneord_d0 <- clean_d0[order(clean_d0$Detector.Name),]

```

<br>
<br>

In the following chunk we want to add columns with information about the experiment, conditions, etc. 

A df with desired columns are first created. 
This column is then bound to the ordered df via cbind() command.

```{r columns with info added - special data D0}

# Create column with important data (Exp_name, Time_point, Replicate, Temp.free_control), which we then want to add to correct dfs

  Exp1_d0 <- rep('Exp1', nrow(geneord_d0))
  Day_0 <- rep('Day 0', nrow(geneord_d0))
  Replicate_0 <- rep(c(2.1, 'NA', 'NA', 'NA', 2.1, 2.1, 2.2, 2.2, 2.2), 8) #uM - Strange order because of how sample name data is loaded into R (as characters)
  Tempfree_control_0 <- rep(c('no', 'yes', 'yes', 'yes', 'no', 'no', 'no', 'no', 'no'), 8) # Same goes for order here as with condition vector.
  
  columns_d0 <- data.frame(Exp1_d0, Day_0, Replicate_0, Tempfree_control_0)
  
  # Add created columns to data frames accordingly
  cbound_d0 <- cbind(columns_d0, geneord_d0)
  
```

<br>
<br>

Here, columns containing information that is of no interest, and the the data (rows) from the negative controls are excluded.
The Exp1 D0 df we are left with contain all necessary data for continued analysis, utilizing whatever R package available. 

The data might need to be modified further to fit the required structure for whatever R package you want to use. 

```{r filtering of data - special data D0}
  
# Create a df with only columns of interest from the cbound_d0 df
  
  col_index_d0 <- c(1:6, 9)

  sorted_d0 <- cbound_d0[,col_index_d0]
  

  # Remove template free control data
  Filtered_d0 <- sorted_d0[sorted_d0$Tempfree_control_0 != 'yes',]
  
  # Assign a new names to the 1-7th columns
  col_names_d0 <- c('Experiment', 'Time point', 'Bioreplica', 'Temp.free control', 'Position', 'Gene', 'Ct')
  names(Filtered_d0)[1:7] <- col_names_d0
    
# Use head() to demonstrate how data looks
# print(head(Filtered_d0))

```

```{r hidden code for head() final - special data D0, echo=FALSE, results='asis'}

# To make the html file look nice: this chunk is added to print the full data, while not showing the actual code in the html file. 
  
  cat("Head of the data frame Filtered_D0:\n")
  cat(capture.output(head(Filtered_d0)), sep="\n")

```
<br>
<br>

### BPT

To get all chiPSC22 data into one df, the average Ct values are calculated and one complete df is created.

```{r create chipsc22 dfs - special data chipsc22}

# Df with chipsc22 data is created

# chiPSC22
  assign(paste("full", '_', 'chiPSC22', sep = ""), data.frame(data_list_special[[1]]))

# Use head() to demonstrate how data looks
# print(head(full_chiPSC22))

```

```{r hidden code for head() - special data chipsc22, echo=FALSE, results='asis'}

# To make the html file look nice: this chunk is added to print the full data, while not showing the actual code in the html file. 

  cat("Head of the data frame full_chiPSC22:\n")
  cat(capture.output(head(full_chiPSC22)), sep="\n")

```

<br>
<br>

In following chunk of code, the data is cleaned. 
Some rows with unnecessary information are excluded. 

```{r clean data from unwanted rows - special data chipsc22}

# Remove rows with unwanted data

  # Created vectors with row numbers that I want to skip to clean data
  skiprow_chipsc22 <- c(7:10, 17:20, 27:30, 37:40, 47:50, 57:60, 67:70, 71) # For chiPSC22 data 


  # Utilize skiprow vector to exclude rows we do not want
  # chiPSC22
  clean_chipsc22 <- full_chiPSC22[-skiprow_chipsc22,]
  
```

<br>
<br>

Here we divide the df 'clean_D0_chipsc22' containing both D0 data and chiPSC22 data - so that a complete df with all chiPSC22 data can be generated.
RSPO3 is put in a separate one, as we only have one set of Ct values of this marker (3 technical replicates).
  
```{r divide D0/chiPSC22 df into useful parts - special data chipsc22}

# Extract chiPSC22 data from D0/chiPSC22 df - create one df without RSPO3 data and one with only RSPO3 data
## Use the one without RSPO3 to calculate average Ct from chiPSC22 qPCR runs

  chipsc_rows <- c(10:12, 22:24, 34:36, 46:48, 70:72, 82:84, 94:96)
  chipsc_wo_rspo3 <- clean_D0_chipsc22[chipsc_rows,]
  chipsc_rspo3 <- clean_D0_chipsc22[58:60,] # Add this one to the final df with averaged Ct:s between runs
  
```

<br>
<br>

Mean Ct values are calculated within dfs and then between so that a common complete df can be created. 
RSPO3 is then added to this (mean Ct values that is).

```{r calculate Ct means within/between chiPSC22 dfs}

# Before averaging chiPSC22 data from 2 separate runs, calculate mean Ct per gene marker in both individual data sets, then average between these

### chiPSC22 from first PCR 

  # Exclude template free controls, because cannot take mean on 'chr' vector otherwise
  NC_rows <- c(4:6, 10:12, 16:18, 22:24, 28:30, 34:36, 40:42)
  clean_chipsc22_wo_NC <- clean_chipsc22[-NC_rows,]


# Mean of Ct per gene marker
  
  n <- 3 # 3 technical replicates we want to average over

  means_list_chipcs <- NULL

  # Calculate means for each gene marker of 3 rows using a for loop
  ## i goes through sequence of length 1 to nrow(df), by n=3 as we have three tech.replicates we want to average
  for (i in seq(1, nrow(clean_chipsc22_wo_NC), by = n)) {
    
    end_i <- min(i + n - 1, nrow(clean_chipsc22_wo_NC))
    means_list_chipcs[[length(means_list_chipcs) + 1]] <- mean(as.numeric(clean_chipsc22_wo_NC[i:end_i, 'Ct']))
    
  }

# Convert the list of means to a matrix or a dataframe
  mean_matrix_chipsc <- do.call(rbind, means_list_chipcs)
  

   
### chiPSC22 from D0/chiPSC22 df

# Mean of Ct per gene marker
  
  n <- 3 # 3 technical replicates we want to average over

  means_list_D0.chipsc <- NULL

  # Calculate means for each gene marker of 3 rows using a for loop
  ## i goes through sequence of length 1 to nrow(df), by n=3 as we have three tech.replicates we want to average
  for (i in seq(1, nrow(chipsc_wo_rspo3), by = n)) {
    
    end_i <- min(i + n - 1, nrow(chipsc_wo_rspo3))
    means_list_D0.chipsc[[length(means_list_D0.chipsc) + 1]] <- mean(as.numeric(chipsc_wo_rspo3[i:end_i, 'Ct']))
    
  }

# Convert the list of means to a matrix or a dataframe
  mean_matrix_D0.chipsc <- do.call(rbind, means_list_D0.chipsc)

  
  
### RSPO3
  mean_RSPO3 <- mean(as.numeric(chipsc_rspo3[, 'Ct']))

  
  
# Calculate mean Ct of chiPSC22 data from both of mean_matrixes generated above
  
  all_chipsc <- data.frame(cbind(mean_matrix_chipsc, mean_matrix_D0.chipsc)) # Except for RSPO3 which is added below ofc
  chipsc_Ct_mean <- rowMeans(all_chipsc, na.rm = TRUE)


```

<br>
<br>

In this chunk we cbind all other column data with the new Ct values
This part might be unecessary, I just liked to make it similar to the other data (Exp1 D 0-13), to have a common structure for further analysis.
Use head() to show final result

```{r merge mean Ct values with rest of the columns - special data chipsc22}

# Extract column data except for Ct values, in order to adhere to the new Ct means
  gene_rows <- seq(1, 37, by = 6)
  common_chipsc_data <- clean_chipsc22[gene_rows, !colnames(clean_chipsc22) %in% 'Ct'] # Exclude Ct column
  RSPO3_data <- clean_D0_chipsc22[clean_D0_chipsc22$Detector.Name == 'RSPO3-Sybr' & clean_D0_chipsc22$Sample.Name == 'D1', !colnames(clean_D0_chipsc22) %in% 'Ct']
    
  
# Merge column data extracted above with calculated Ct mean values
  # All chipsc wo RSPO3
  merged_chips <- cbind(common_chipsc_data[,1:4], chipsc_Ct_mean, common_chipsc_data[,5:32])
  # RSPO3
  merged_RSPO3 <- cbind(RSPO3_data[1:4], mean_RSPO3, RSPO3_data[5:32])
  colnames(merged_RSPO3) <- names(merged_chips)
  
  
# rbind() merged_chipsc and merged_RSPO3 to get all mean Ct values from chiPSC22 in one df
  complete_chipsc_mean <- rbind(merged_chips, merged_RSPO3)

  
# Use head() to demonstrate how data looks
# print(head(complete_chipsc_mean))

```

```{r hidden code for head() mean - special data chipsc22, echo=FALSE, results='asis'}

# To make the html file look nice: this chunk is added to print the full data, while not showing the actual code in the html file. 
  
  cat("Head of the data frame with mean Ct values and all columns:\n")
  cat(capture.output(head(complete_chipsc_mean)), sep="\n")

```

<br>
<br>

Here, data is ordered after gene markers (according to how they are structured on the plate). 
  
```{r order after genes - special data chipsc22}
  
# Order data frames after desired gene marker order

  # First, convert the 'gene' column to a factor with levels in the desired order
  complete_chipsc_mean$Detector.Name <- factor(complete_chipsc_mean$Detector.Name, 
                                     levels = c('Actb-SYBR', 'GAPDH-Syber', 'SOX17-Sybr', 'RSPO3-Sybr', 'SPINK2-Sybr', 'HOXA9-Sybr', 'Runx1', 'Gata2-Sybr'))

  # Order data accordingly
  geneord_chipsc22 <- complete_chipsc_mean[order(complete_chipsc_mean$Detector.Name),]
  
```

<br>
<br>

Here, columns containing information that is of no interest are excluded.
The chiPSC22 df we are left with contain all necessary data for continued analysis, utilizing whatever R package available. 

The data might need to be modified further to fit the required structure for whatever R package you want to use. 

```{r columns with info added & filtering of data - special data chipsc22}

# Create column with important data (sample, time point, position), which we then want to add to correct the chiPSC22 df
  Sample <- rep('chiPSC22', nrow(geneord_chipsc22))
  Day_chipsc <- rep('Day -1', nrow(geneord_chipsc22))
  Row_position <- c('F', 'G', 'A', 'D_rspo3', 'B', 'C', 'D_runx1', 'E')
  
  columns_chipsc <- data.frame(Sample, Day_chipsc, Row_position)
  
  # Add created columns to data frames accordingly
  cbound_chipsc22 <- cbind(columns_chipsc, geneord_chipsc22)
  
  
# Create a df with only columns of interest from the cbound_chipsc22 df
  
  col_index_chipsc <- c(1:3, 5, 8)
  
  Filtered_chipsc22 <- cbound_chipsc22[,col_index_chipsc]
  
  # Assign a new names to the 1-7th columns
  col_names_chipsc <- c('Sample', 'Time point', 'Row position', 'Gene', 'Mean Ct')
  names(Filtered_chipsc22)[1:5] <- col_names_chipsc
  
  
# Use head() to demonstrate how data looks
# print(head(Filtered_chipsc22))
  
```

```{r hidden code for head() final - special data chipsc22, echo=FALSE, results='asis'}

# To make the html file look nice: this chunk is added to print the full data, while not showing the actual code in the html file. 
  
  cat("Head of the data frame Filtered_chipsc22:\n")
  cat(capture.output(head(Filtered_chipsc22)), sep="\n")

```

<br>

### Save the filtered data for further data analysis

Save the filtered dfs, so they can be accessed by the Data Analysis Notebook.

```{r save filtered data - all data}

# Create one complete list with all dfs, name them accordingly otherwise will be ordered as [[i]] in list
complete_filtered_list <- c(list(Filtered_d0), Filtered_list_common, list(Filtered_chipsc22))
names(complete_filtered_list) <- c('Filtered_d0', 'Filtered_d3.1', 'Filtered_d3.2', 'Filtered_d8.1', 'Filtered_d8.2', 
                                   'Filtered_d13.1', 'Filtered_d13.2', 'Filtered_chipsc22')

# Save to computer
save(complete_filtered_list, file = "all_filtered.RData")

```


